---
title: "state_profile_florida"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PPP State Profile

Your assignment is to explore a dataset of PPP loan applications for a particular state and to answer a series of questions that will help you summarize that state's loan applications. You will need to write R code to answer those questions. You also will bring in other data, including Census information, and you will be making graphics, including maps, to illustrate your findings.

The deliverable will be this R Markdown notebook and a data folder that you will receive with a state's loan application data. Place any other data you are using to answer the questions in the same data folder.

**Task** Create a codeblock and load appropriate packages and settings for this assignment.

* A Census API key
```{r echo=FALSE}
census_api_key("b6d29189eb33214a9a0b6a811a718e9ec0543cb6")
```

## Load libraries and establish settings
```{r}
library(tidyverse)
library(tidycensus)
library(tigris)
library(janitor)
library(lubridate)
library(ggrepel)
library(sf)

```

## Initial Exploration

**Q1.**. Write R code that generates some basic descriptive statistics that allows you to describe the applications from your state. This code should produce results that describe the data in different ways: how many applications there are, along with ways to understand the typical and most frequent values for columns you find interesting or newsworthy. You should produce at least five (5) descriptive statistics and write up a summary of the data using them.

**A1.** There are 983,362 applications in Florida for PPP, according to the dataset.
Miami-Dade county got the highest number of loans with 222,558. 22 loans were recorded with NA as their county name. 
Miami-Dade also had the highest number of jobs retained with 1,094,393. About 11 counties got the max loan amount of 10,000,000. According to census reporter, Miami-Dade county has a population of 2,705,528, and therefore, almost half the population had their jobs retained which was interesting.

```{r}
florida_ppp <- read_csv("C:/Users/lolod/Downloads/florida_ppp.csv/florida_ppp.csv")
florida_ppp <- florida_ppp %>% mutate(city = str_to_title(city))

#The number of PPP loan applications in Florida

florida_ppp %>% summarise(total_ppp_loans = n())

#The Florida county which got more loans

florida_ppp %>%
  group_by(project_county_name)%>%
  summarise(
    count_loans = n()
  ) %>% arrange(desc(count_loans))

#Which records had the county name as NA
florida_ppp %>% filter(is.na(project_county_name))

#Finding the mean, median, min and max loan amounts
florida_ppp %>%
  group_by(project_county_name) %>%
  summarise(
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))

#Number of jobs retained

florida_ppp %>%
  group_by(project_county_name)%>%
  summarise(
    total_jobs = sum(jobs_retained)
    ) %>% arrange(desc(total_jobs))

```


## Geographic Analysis

**Q2.** Write R code that examines geographic patterns for PPP loans in your state, using Census population information to calculate a per-capita figure for the state and counties and zip codes. Then, make a county map using ggplot showing the per-capita data and a zip code map showing the difference from the statewide per-capita figure. Describe the most interesting or newsworthy findings based on your exploration.

**A2.** Florida has 4704 loans per 100,000 people. Monroe county has the highest number of loans per capita for a population of 75,798. In the statewide per capita map, most of the state has about 1000 loans per zip code.This seems to be an average high. 
There's a larger difference between the loans in the north west and south of the state in the state versus county maps.


```{r}
#state

fl_state <-  get_acs(geography = "state", variables = "B01001_001", year = 2019, geometry = TRUE) %>% filter(NAME == "Florida")

fl_ppp_loans <- florida_ppp %>% 
  summarise(count=n())
state_per_capita <- fl_state %>% 
  mutate(
    per_capita = ((fl_ppp_loans$count / fl_state$estimate)*100000)
    
  )

#county

fl_county <- get_acs(geography = "county", variables = c(population = "B01001_001"), 
              state = "FL", geometry = TRUE)

  fl_county_clean <- fl_county %>% 
  rename(COUNTY = NAME) %>%
  mutate(COUNTY = toupper(str_remove_all(COUNTY,", Florida|County"))) %>%
  mutate(COUNTY = str_trim(COUNTY,side="both")) %>%
  rename(population = estimate) %>%
  select(COUNTY, population, geometry)
  
  fl_county_loans <- florida_ppp %>% 
  group_by(project_county_name) %>% 
  summarise(count_loans = n()) %>% 
  rename("COUNTY" = "project_county_name")
  
 county_per_capita<- merge(fl_county_clean, fl_county_loans, by=c("COUNTY"), all.COUNTY=FALSE) %>% 
  mutate(loans_per_capita = ((count_loans/population)* 100000)) %>%
   arrange(desc(loans_per_capita))

ggplot() + 
  geom_sf(data=county_per_capita, aes(fill=loans_per_capita)) + 
  theme_minimal()+
  scale_fill_viridis_b(option="magma",trans = "log") 

#zipcode

fl_zip <- get_acs(geography = "zcta", variables = c(population = "B01001_001"), state = "FL", geometry = TRUE)

florida_ppp <- florida_ppp %>% 
  mutate(zip5 = str_sub(zip, start=1L, end=5L))

fl_zip_clean <- florida_ppp %>% 
  mutate(project_zip = str_sub(project_zip, start = 1L, end = 5L)) %>% 
  rename(GEOID = project_zip) %>% 
  group_by(GEOID) %>% 
  summarise(total = n())

zip_per_capita <- fl_zip %>% 
  left_join(fl_zip_clean, by = "GEOID") %>% 
  mutate(per_capita = ((total/estimate) *100000),
         per_capita_diff = state_per_capita$per_capita - per_capita)

 
state_map <- counties() %>% 
  filter(STATEFP == 12)
ggplot() + 
  geom_sf(data=state_map, fill = "white")+
  geom_sf(data=zip_per_capita, aes(fill=per_capita_diff)) + 
  theme_minimal()+
  scale_fill_viridis_b(option="magma",trans = "log") 


```


## Lender Analysis

**Q3.** Write R code to examine which lenders had the most approved applications in your state (and include summary statistics such as total amount and average loan amount). Generate dataframes that show the number and total amount of all loans and undisbursed loans per lender. For those lenders who had any undisbursed loans, make a graphic showing the relationship between the total amount of loans and the total undisbursed amount. Describe the most noticeable outlier lenders on that graphic below.

**A3.** By comparing the total amount of all loans per lender to the total undisbursed loan amounts, we got four lenders which had undisbursed loan amounts in Florida. These four lenders are Blue Ridge Bank, National Association; The Bancorp Bank; Fountainhead SBF LLC and Centennial Bank.

https://datawrapper.dwcdn.net/ytxY2/1/ 

```{r}
fl_undisbursed_amount <- florida_ppp %>% 
  select(name, amount, lender, undisbursed_amount, zip5, date_approved) %>%
  arrange(desc(undisbursed_amount))

lenders_undisbursed <- fl_undisbursed_amount %>% 
  group_by(lender) %>% 
  summarise(total_undisbursed_amount = sum(undisbursed_amount),
            total_amount = sum(amount)) %>%
  arrange(desc(total_undisbursed_amount)) %>% 
  head(10)

write_csv(lenders_undisbursed, "fl_lenders_undisbursed.csv")


```


## Industry Analysis

**Q4.** Write R code that examines industry patterns for PPP loans in your state, using the NAICS codes from the PPP data as a starting point. Generate statewide and county industry totals, then join that with 2018-19 data from the [Census County Business Patterns survey](https://www2.census.gov/programs-surveys/cbp/datasets/2019/cbp19co.zip) using 6-digit NAICS codes. The documentation explaining that data can be found here: https://www2.census.gov/programs-surveys/cbp/technical-documentation/records-layouts/2018_record_layouts/county-layout-2018.txt. To do this, you will need to add FIPS codes to your PPP dataset.

Does the distribution of PPP applications by the top 10 industries (by number of applications) roughly match the number of businesses reported in the Census data? Does it roughly match if you remove self-employed individuals and sole proprietorships from the PPP totals? Write up a summary of what you've found and whether there are potentially newsworthy patterns in the data.

Create a county-level map showing the differences between the PPP applications and the Census business data for one of the top 10 industry codes. You can do this either using ggplot or Datawrapper (if the latter, publish that map and include the URL below).

**A4.** The distribution of PPP applications by industries roughly match the number of businesses reported in the census data but only for some industries. One industry that stood out was the taxi service industry. The taxi service industry came in second in top 10 industries which applied for PPP loans in Florida state. But it was also the industry with the least number of establishments. 

https://datawrapper.dwcdn.net/KNESo/1/
https://datawrapper.dwcdn.net/Tu47P/1/

```{r}
cbp_19_co <- read_csv("cbp19co.csv")

florida_cbp <- cbp_19_co %>% 
  filter(fipstate == 12)

naics_codes <- read_csv("naics_codes.csv")

florida_ppp_naics <- florida_ppp %>% 
  inner_join(naics_codes, by="naics_code")%>% 
  select(name, naics_code, title, project_county_name)

county_fips <- read_csv("county_fips.csv") 

fl_county_fips <- county_fips %>% 
  filter(state == "FL")

florida_ppp_naics_fips <- florida_ppp_naics %>% 
  inner_join(fl_county_fips, by=c("project_county_name" = "match_county")) %>% 
  select(-state, -state_code, -state_name, -county_code, -county)

#top ten industries by state
fl_ppp_naics_fips_state <- florida_ppp_naics_fips %>% 
  group_by(naics_code) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total)) %>% 
  head(10)

#top ten industries by county
fl_ppp_naics_fips_county <- florida_ppp_naics_fips %>% 
  group_by(fips_code) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total)) %>% 
  head(10)

fl_ppp_naics_fips_state <- fl_ppp_naics_fips_state %>% 
  left_join(naics_codes, by=c("naics_code")) 

florida_cbp <- florida_cbp %>% rename(naics_code=naics)

fl_naics_est<- merge(florida_cbp,fl_ppp_naics_fips_state,by=c("naics_code"), all.naics_code = FALSE)

fl_naics_est<-fl_naics_est %>% 
  group_by(naics_code) %>% 
  summarise(
    est=sum(est)
  ) %>% mutate(naics_code= as.numeric(naics_code)) 

fl_ppp_naics_fips_state<-fl_ppp_naics_fips_state %>%
  inner_join(fl_naics_est, by=c("naics_code"))
#Not all of the PPP applications numbers match the number of businesses.

# Using naics code 485310 for the map because it has the greatest difference between the loan applicatons and the number of businesses.
fl_naics_485310 <- florida_ppp_naics_fips %>% 
  filter(naics_code == "485310") %>% 
  group_by(fips_code) %>% 
  summarise(total = n())

fl_est_cbp<- florida_cbp %>% 
  filter(naics_code == "485310")

fl_est_cbp$fips <- paste0(fl_est_cbp$fipstate, fl_est_cbp$fipscty)

fl_est_cbp <- fl_est_cbp %>% 
  select(fips, est)


write_csv(fl_naics_485310, "fl_naics_485310.csv")
write_csv(fl_est_cbp, "fl_cbp_485310.csv")

```


## Summary

**Q5.** What is the most promising story idea that you have found during this exploration? How would you approach reporting it out and what other information/data would you need to do that? What would be the minimum story you could get from this, and what would be the maximum story?

**A5.**
```{r}






```