---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.



##Jan.27 lab 
setwd() is the command for setting working directory

```{r}
setwd("~/GitHub/data_journalism_2022_spring") 

```

```{r}
install.packages("tidyverse")

```
```{r}
library(tidyverse)
```
## Command to change datatype
 column name = as.data type(column name)
 Ex: id=as.character(id)
##Run the following codeblock to find the largest and smallest loans in the data, by using summarise() without grouping.

## 
rename() command is part of tidyverse and not janitor.
janitor commands:
clean_names() command cleans up the names
glimpse() shows the columns and datatypes
distinct() will get rid of duplicate records
get_dupes() will display the duplicated data
str_sub() will create a substring that can restrict the size of the values. It's used along mutate function.
str_to_title() will make values titlecase (not lower or upper)
city== "smth", "smthh"
TRUE~city -- true for all city values except for smth smth.

## Follow the format of:
filter()
mutate()
group_by()
summarize()
arrange()

## use filter(str_detect(column_name, "name | name")) to filter out specific values
## bind_rows(list(datasetname,datasetname, datasetname)) is used to combine diff datasets to form a single one with data from all 3 datasets.
## left_join() adds a separate data set to the main data set. 
