---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
census_api_key("b6d29189eb33214a9a0b6a811a718e9ec0543cb6") 
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
ACS <- load_variables(2019, "acs5", cache=TRUE )

```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
#Write code here to read the RDS file
ppp <- read_rds("data/lab_08.rds")
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first in your list of ACS variables.
**A1** Baltimore city 

```{r}

acs19 <- get_acs(geography = "county", state="MD", variables = "B01002_001", year = 2019)
acs19 %>% arrange(estimate)

```

**Q2** Which Maryland counties have a median age of at least 50? Write code to isolate just those jurisdictions.
**A2** Talbot and Worcester counties have a median age of at least 50.

```{r}
acs19 %>% filter (estimate >= 50)



```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? You'll need one dataframe with county-level data on the number of 5-17 year-olds that speak Spanish at home and a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3** Prince George's and Montgomery counties have the highest percent of 5-17 year-olds that speak Spanish.

```{r}
acs19_spanish <- get_acs(geography = "county", state="MD", variables = "B16004_002", year = 2019)

acs19_total <- get_acs(geography = "county", state="MD", variables = "B16004_004", year = 2019)

acs19_percent <- acs19_spanish %>% inner_join(acs19_total, by=c("GEOID", "NAME"))

acs19_percent %>% rename(estimate_total=estimate.x, estimate_spanish=estimate.y) %>%
  select(-variable.x, -variable.y, -moe.x, -moe.y) %>%
  mutate(percentage = ((estimate_spanish/estimate_total)*100)) %>%
  arrange(desc(percentage))




```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4** The counties with largest percentage of 5-17 year-olds that speak Spanish at home are Prince George's, Montgomery, Caroline, Frederick and Talbot. Prince George's, Montgomery counties have a diverse population and the results make sense.Frederick and Talbot counties are predominantly white populated counties.
Garrett County has only 0.7% of Spanish speakers in 5-17 year olds. It has a 96% white population and the results make sense.


**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5** ZIP 60620 has the most loans.95.6% is the percentage of Black population in the ZIP 60620.

```{r}
ppp <- ppp %>%
  mutate(zip5 = str_sub(zip, 1, 5))
highest_ppp <- ppp %>% group_by(zip5) %>% 
  summarize(
    count_loans=n()
  ) %>% 
arrange(desc(count_loans))

acs_zipcodes <- get_acs(geography="zcta", variables = "B01001_001")
acs_black <- get_acs(geography="zcta", variables = "B01001B_001")

blackpercent <- acs_zipcodes %>% inner_join(acs_black, by=c("GEOID", "NAME"))
blackpercent %>% rename(estimate_total=estimate.x, estimate_black=estimate.y) %>%
  select(-variable.x, -variable.y, -moe.x, -moe.y) %>%
  filter(GEOID == "60620") %>% 
  mutate(percentage = ((estimate_black/estimate_total)*100)) %>%
  arrange(desc(percentage))





```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed applications in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6** The top 10 zip codes with the most removed applications in applications in terms of Black population belong to Chicago, Illinois and Atlanta, Georgia. Both these places have a fairly high Black population according to census reporter.


**Q7** What other columns from the PPP data could be useful in reporting on what happened to these removed loans?
**A7** Gender and Lenders columns from the PPP data could be useful in reporting on these removed loans. Did these removed loans pertain more to a certain gender? Or did one or more lenders reject loans more often than others? If so, on what basis did they deny these loans?
